---
title: "Academics preprocessing"
author: "kaiiii"
date: "2026-09-08"
output: html_document
---

A previous iteration of data imputation conducted EM using all datapoints in the full TeacherGAIA dataset. This present script creates imputed academic results **within** the final dataset for the TeacherGAIA Info-seek project.


** Key outputs **
df_nonimputed - 156 complete case data
df_imputed - 250 academic data points acquired using the EM algorithm


```{r}
library(readxl)
library(writexl)
library(tidyverse)
library(Amelia)
```

```{r}
root_dir <- "~/tgaia"
```

# Set up & cleaning 
```{r import relevant datasets}
df_academics <- read_excel(file.path(root_dir, "processed_data/academic_outcomes.xlsx"))

df_selected <- read_excel(file.path(root_dir, "processed_data/conversation_logs.xlsx"))
```

```{r check duplicate academic outcomes from each student}
# keep first response if duplicates
dup_ids_acad <- df_academics %>% count(studId) %>% filter(n > 1) %>% pull(studId)

dupes_first_acad <- df_academics %>%
  filter(studId %in% dup_ids_acad) %>%
  arrange(studId) %>%
  group_by(studId) %>% slice(1) %>% ungroup()

df_academics <- df_academics %>%
  filter(!studId %in% dup_ids_acad) %>%
  bind_rows(dupes_first_acad)
```


```{r get a list of studIds from the selected modules}
df_selected %>%
  dplyr::select(studId) %>%
  distinct() %>%
  pull() -> selected_studIds

df_student_info <- df_selected %>% 
  mutate(dataset = as.factor(dataset)) %>% 
  group_by(studId) %>% 
  summarise(dataset = first(dataset),
            schoolLevel = first(schoolLevel))
```

# Create a nonimputed dataset
```{r non imputed dataset}
df_nonimputed <- df_academics %>% 
  filter(studId %in% selected_studIds) %>% 
  rename(pre_tgaia = assignment1_NOAI,
         post_tgaia = assignment2_AI) %>%
  mutate(pre_tgaia = as.numeric(pre_tgaia),
         post_tgaia = as.numeric(post_tgaia)) %>% 
  left_join(df_student_info, by = "studId")

colSums(is.na(df_nonimputed))
mean(is.na(df_nonimputed$pre_tgaia)) * 100
mean(is.na(df_nonimputed$post_tgaia)) * 100


# standardise scores within each dataset
df_nonimputed <- df_nonimputed %>%
  group_by(dataset) %>%
  mutate(pre_tgaia_z = scale(pre_tgaia)[, 1],
         post_tgaia_z = scale(post_tgaia)[, 1]) %>%
  ungroup() %>% 
  dplyr::select(-pre_tgaia, -post_tgaia, -dataset, -schoolLevel) %>% 
  drop_na

```



# Create an imputed dataset

```{r prep a df for imputation}
df_academics_clean <- df_academics %>% 
  filter(studId %in% selected_studIds) %>% 
  rename(pre_tgaia = assignment1_NOAI,
         post_tgaia = assignment2_AI) %>% 
  mutate(across(c(pre_tgaia, post_tgaia), 
                ~ifelse(. == "NaN", NA, .))) %>%  
  mutate(pre_tgaia = as.numeric(pre_tgaia),
         post_tgaia = as.numeric(post_tgaia)) %>% 
  left_join(df_student_info, by = "studId")

df_input <- df_academics_clean %>%
  select(studId,pre_tgaia, post_tgaia) 
```


```{r check missingness}
colSums(is.na(df_input))
```


```{r expectation maximisation}
df_output <- amelia(as.data.frame(df_input, stringsAsFactors = FALSE), 
                                m = 1, 
                                seed = 123,
                                idvars = "studId")$imputations$imp1
colSums(is.na(df_output))
```

```{r standardise scores within each dataset}
df_imputed <- df_output %>%
  left_join(df_student_info, by = "studId") %>%
  group_by(dataset) %>%
  mutate(pre_tgaia_z = scale(pre_tgaia)[, 1],
         post_tgaia_z = scale(post_tgaia)[, 1]) %>%
  ungroup() %>% 
  select(-pre_tgaia, -post_tgaia, -dataset, -schoolLevel) %>% 
  drop_na
```

# Outputs
```{r}
write_xlsx(df_nonimputed, file.path(root_dir, "processed_data/acads_nonimputed.xlsx"))
write_xlsx(df_imputed, file.path(root_dir, "processed_data/acads_imputed.xlsx"))
```

