---
title: "TeacherGAIA main analysis script"
author: "kaiiii"
date: "2026-09-08"
output: 
  html_document:
    toc: TRUE
    toc_depth: 2
---

Revised statistical analysis script after receiving reviewers' comments in Sept 2026. 

We investigated how trait curiosity related to distinct information seeking profiles. Information seeking was characterised at prompt-level using three classifications (agency, scope, intent) and two semantic distance measures (mean SD and mean TD). 

Then, we examined if these profiles led to tangible learning after engagement in genAI-facilitated learning modules.

Queries and feedback may be directed to the corresponding author jiahou.poh[at]nus.edu.sg.

# Set up
Required packages
```{r required packages, include = FALSE}
pkgs = c("readxl", "ggeffects", "grid", "gridExtra", "car", "psych", "patchwork", "tidyverse", "svglite", "emmeans", "interactions", "tibble", "tidyr", "lme4", "lmerTest", "broom", "officer", "flextable", "effectsize")

inst = lapply(pkgs, library, character.only=TRUE)

set.seed(555)
```
```{r set root dir, include = FALSE}
root_dir <-"~/tgaia/"
```
```{r set up type iii contrasts}
options(contrasts = c("contr.sum", "contr.poly"))
```


Import data
```{r data importing}
# (1) Selected observations
data_chatlogs <- readxl::read_xlsx(file.path(root_dir,'processed_data/conversation_logs.xlsx')) %>% 
  mutate(dataset = as.factor(dataset),
         studId = as.factor(studId))

# (2) Agency (Active vs. Passive)
data_query_intent <- readxl::read_xlsx(file.path(root_dir,'processed_data/agency_majority.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
    query_type = case_when(
      final_vote == "Active Query" ~ 1L, # indicates an information or task request has been made
      final_vote == "Passive Response" ~ 0L,
      TRUE ~ NA_integer_
    )
  ) %>% 
  dplyr::select(id, studId, query_type)


# (3) Scope (Open vs. Closed)
data_prompt_scope <- readxl::read_xlsx(file.path(root_dir,'processed_data/scope_majority.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
    prompt_scope = case_when(
      final_vote == "Close-ended" ~ 1L, # indicates a narrow, close-ended query
      final_vote == "Open-ended" ~ 2L, # indicates a broad, open-ended query
      final_vote == "Not Applicable" ~ 3L,
      TRUE ~ NA_integer_
    )
  )%>% 
  dplyr::select(id, studId, prompt_scope)

# (4) Intent categories
data_cognitive_intent <- readxl::read_xlsx(file.path(root_dir,'processed_data/intent_majority.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
    cognitive_intent = case_when(
      final_vote == "Remembering" ~ 1L,
      final_vote == "Understanding" ~ 2L,
      final_vote == "Applying" ~ 3L,
      final_vote == "Analysing" ~ 4L,
      final_vote == "Evaluating" ~ 5L,
      final_vote == "Creating"  ~ 6L,
      final_vote == "Irrelevant" ~ 7L,
      TRUE ~ NA_integer_
    )
  )%>% 
  dplyr::select(id, studId, cognitive_intent)

# (5) Semantic distance metrics
data_semantic_distance <- readxl::read_xlsx(file.path(root_dir,'processed_data/semantic_dist.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
    dataset = as.factor(dataset)
  )

# (6) 5DCR
data_curiosity_measures <- readxl::read_xlsx(file.path(root_dir,'processed_data/replication_5dcr.xlsx')) %>%
  mutate(studId = as.factor(studId)) %>%
  drop_na(studId)

# (7) Academic outcomes 
data_academic_outcomes_imputed <- readxl::read_xlsx(file.path(root_dir, 'processed_data/acads_imputed.xlsx')) %>% 
  mutate(studId = as.factor(studId))

data_academics_outcomes_nonimputed <- readxl::read_xlsx(file.path(root_dir,'processed_data/acads_nonimputed.xlsx')) %>% 
  mutate(studId = as.factor(studId))
```

# Student-level dfs for analyses

## Trait curiosity (5DCR)
```{r curiosity df}
df_5dcr <- data_curiosity_measures
```

# Student-level prompting behaviours
```{r define logit function}
logit <- function(p) {
  p <- pmin(pmax(p, 1e-6), 1 - 1e-6) # small offset to avoid undefined values
  log(p / (1 - p))} # continuity corr
```

```{r agency scope intent semantic distance dfs}
data_query_intent       <- data_query_intent       %>% mutate(id = as.character(id))
data_prompt_scope       <- data_prompt_scope       %>% mutate(id = as.character(id))
data_cognitive_intent   <- data_cognitive_intent   %>% mutate(id = as.character(id))
data_semantic_distance  <- data_semantic_distance  %>% mutate(id = as.character(id))


#1) Agency (Active vs. Passive)
df_query_intent <- data_query_intent %>%
  group_by(studId) %>%
  summarise(
    n_active_query      = sum(query_type == 1, na.rm = TRUE),
    n_passive_response  = sum(query_type == 0, na.rm = TRUE),
    n_total             = n_active_query + n_passive_response,
    .groups = "drop"
  ) %>%
  
  mutate(
    prop_active = (n_active_query) / (n_total),
    prop_active_adj = (n_active_query + 0.5) / (n_total + 1),
    logit_active = logit(prop_active_adj),
    c_logit_active = as.numeric(scale(logit_active, scale = TRUE))) 
  

# 2) Scope (Open vs. Closed)
df_prompt_scope <- data_prompt_scope %>%
  group_by(studId) %>%
  summarise(
    n_open  = sum(prompt_scope == 2, na.rm = TRUE),
    n_close = sum(prompt_scope == 1, na.rm = TRUE),
    n_na    = sum(prompt_scope == 3, na.rm = TRUE),
    n_openclose = n_open + n_close,
    n_total = n(),
    .groups = "drop"
  ) %>%
  mutate(
    prop_open = (n_open) / (n_openclose),
    prop_open_adj = (n_open + 0.5) / (n_openclose + 1),
    logit_open = logit(prop_open_adj),
    c_logit_open = as.numeric(scale(logit_open, scale = TRUE)))

# 3) Intent Categories
df_cognitive_intent <- data_cognitive_intent %>%
  mutate(
    is_high = as.integer(cognitive_intent %in% c(4,5,6)),
    is_low  = as.integer(cognitive_intent %in% c(1,2,3)),
    is_na = as.integer(cognitive_intent == 7)
  ) %>%
  group_by(studId) %>%
  summarise(
    n_high = sum(is_high, na.rm = TRUE),
    n_low  = sum(is_low,  na.rm = TRUE),
    n_is_not_na = n_high + n_low,
    n_is_uncat = sum(is_na),
    n_total = n(),
      .groups = "drop"
  ) %>%
  mutate(
    prop_high = (n_high) / (n_is_not_na),
    prop_high_adj = (n_high + 0.5) / (n_is_not_na + 1),
    logit_high = logit(prop_high_adj),
    c_logit_high = as.numeric(scale(logit_high, scale = TRUE)))


# 4) Creating
df_offloading <- data_cognitive_intent %>%
  group_by(studId) %>%
  summarise(
    n_creating = sum(cognitive_intent == 6),
    n_is_not_na = sum(cognitive_intent < 7),
    n_total = n(),
      .groups = "drop"
  ) %>%
  mutate(
    prop_off_adj = (n_creating + 0.5) / (n_is_not_na + 1),
    logit_off = logit(prop_off_adj),
    c_logit_off = as.numeric(scale(logit_off, scale = TRUE))) 

# 5) Semantic distance measures
df_semantic_measures <- data_semantic_distance %>%
  group_by(studId) %>%
  summarise(n_total = n(),
            mean_p2p = mean(p2p, na.rm = TRUE),
            mean_sys = mean(sys)) %>% 
  mutate(mean_p2p_dist = 1-mean_p2p, # reverse-scored so higher scores reflect greater dissimilarity 
         mean_sys_dist = 1-mean_sys) %>% 
  dplyr::select(studId, mean_p2p_dist, mean_sys_dist, n_total) %>%  
  mutate(
    mean_p2p_dist_z = as.numeric(scale(mean_p2p_dist, scale = TRUE)),
    mean_sys_dist_z = as.numeric(scale(mean_sys_dist, scale = TRUE))
  )

```

## Covariates and 5CDR 
```{r create a df with the 5dcr values but also with covariates}
df_covariates <- data_chatlogs %>% 
  group_by(studId) %>% 
  mutate(n_total = n(),
        log_total = log1p(n_total)) %>% 
  ungroup() %>% 
  mutate(c_log_total = as.numeric(scale(log_total, scale = TRUE))) %>% 
  dplyr::select(studId, dataset, n_total, log_total, c_log_total) %>%
  group_by(studId) %>% 
  summarise(
    dataset = first(dataset),
    n_total = first(n_total),
    log_total = first(log_total),
    c_log_total = first(c_log_total)
  )

df_curiosity_with_covars <- df_5dcr %>%
  dplyr::select(studId, joyExp, depSen, strTol) %>% 
  left_join(df_covariates, by = "studId")

n_distinct(df_curiosity_with_covars %>% drop_na()) # how many intersect?
```

# PCA
```{r pca prep + components selection}
# Prepare clean dataframe for PCA
df_pca_ready <- df_covariates %>% dplyr::select(studId) %>% 
  left_join(df_prompt_scope %>% dplyr::select(studId, c_logit_open), by = "studId") %>%
  left_join(df_query_intent %>% dplyr::select(studId, c_logit_active), by = "studId") %>%
  left_join(df_cognitive_intent %>% dplyr::select(studId,c_logit_high), by = "studId") %>%
  left_join(df_offloading%>% dplyr::select(studId, c_logit_off), by = "studId") %>%
  left_join(df_semantic_measures %>% dplyr::select(studId, mean_p2p_dist_z, mean_sys_dist_z), by = "studId") %>%
  drop_na(c_logit_active, c_logit_open, c_logit_high, mean_p2p_dist_z, mean_sys_dist_z) # remove NAs for PCA

print(paste("no. of students for PCA:", length(unique(df_pca_ready$studId))))

# Select and scale the 5 indicators
pca_input_clean <- df_pca_ready %>%
  dplyr::select(c_logit_active, c_logit_open, c_logit_high, mean_p2p_dist_z, mean_sys_dist_z) %>%
  as.data.frame()

# Parallel Analysis to select numbers of PC
parallel_results <- fa.parallel(pca_input_clean, 
                                fa = "pc", 
                                n.iter = 100, 
                                show.legend = TRUE, 
                                main = "Scree Plot with Parallel Analysis")

# save the plot

```

```{r eigenvalues}
# Kaiser Criterion (Eigenvalues > 1.0)
actual_eigenvalues <- eigen(cor(pca_input_clean))$values
print("Actual Eigenvalues:")
print(round(actual_eigenvalues, 3))
```
Kaiser criterion and elbow method suggested a 2-component solution.

```{r extract PCs using selected solution}
final_pca_model <- principal(pca_input_clean, nfactors = 2, rotate = "varimax")
print(final_pca_model$loadings, cutoff = 0.3)
```
```{r reverse Comp1 for interpretability}
# Extract scores back into dataframe
df_pca_ready$Comp1 <- final_pca_model$scores[, "RC1"]
df_pca_ready$Comp2 <- final_pca_model$scores[, "RC2"]
df_pca_ready$Comp1 <- -final_pca_model$scores[, "RC1"] # reverse score PC1 for greater interpretability
```

# Statistical analyses

## How does trait curiosity relate to overall prompt volume?
```{r}
m0_data_5dcr_promptvol <- df_curiosity_with_covars %>%
  drop_na()

m0_fit_5dcr_promptvol <- lm(n_total ~ depSen * joyExp * strTol + dataset, data = m0_data_5dcr_promptvol)
summary(m0_fit_5dcr_promptvol)
Anova(m0_fit_5dcr_promptvol, type = 3)
```


## How does trait curiosity predict PCA components? 
```{r prep data for 5DCR ~ PCA}
m1_data_5dcr_pc <- df_pca_ready %>%
  left_join(df_curiosity_with_covars, by = "studId") %>% 
  drop_na()

print(paste("no. of students for 5DCR ~ PCA:", length(unique(m1_data_5dcr_pc$studId))))
```


```{r Does 5DCR predict PC1?}
m1a_fit_5dcr_comp1 <- lm(Comp1 ~ depSen * joyExp * strTol + c_log_total + dataset, data = m1_data_5dcr_pc)
summary(m1a_fit_5dcr_comp1)
Anova(m1a_fit_5dcr_comp1, type = 3)
eta_squared(car::Anova(m1a_fit_5dcr_comp1, type = 3), partial = TRUE)
confint.default(m1a_fit_5dcr_comp1, level = 0.95)
```

```{r Does 5DCR predict PC2?}
m1b_fit_5dcr_comp2 <- lm(Comp2 ~ depSen * joyExp * strTol +  c_log_total + dataset, data = m1_data_5dcr_pc)
summary(m1b_fit_5dcr_comp2)
Anova(m1b_fit_5dcr_comp2, type =3)

```

We found interaction effects of JE x DS and JE x ST on PC1.

Simple slopes analyses:
```{r prep for PC1 simple slopes}
m_comp1_je_mean  <- mean(m1_data_5dcr_pc$joyExp, na.rm = TRUE)
m_comp1_je_sd <- sd(m1_data_5dcr_pc$joyExp, na.rm = TRUE)

m_comp1_ds_mean  <- mean(m1_data_5dcr_pc$depSen, na.rm = TRUE)
m_comp1_ds_sd <- sd(m1_data_5dcr_pc$depSen, na.rm = TRUE)

m_comp1_st_mean  <- mean(m1_data_5dcr_pc$strTol, na.rm = TRUE)
m_comp1_st_sd <- sd(m1_data_5dcr_pc$strTol, na.rm = TRUE)
```


```{r simple slopes for JE x DS interaction}
ss_comp1_je_ds <- emtrends(
  m1a_fit_5dcr_comp1,
  ~ joyExp,
  var = "depSen",
  at = list(
    joyExp = c(m_comp1_je_mean - m_comp1_je_sd, m_comp1_je_mean + m_comp1_je_sd),
    strTol = m_comp1_st_mean   # hold ST at mean
  )
)

summary(ss_comp1_je_ds, infer = c(TRUE, TRUE))
pairs(ss_comp1_je_ds)
```
When JE is low, DS is positively associated with greater epistemic focus

```{r otherview for perspective}
ss_comp1_je_ds <- emtrends(
  m1a_fit_5dcr_comp1,
  ~ depSen,
  var = "joyExp",
  at = list(
    depSen = c(m_comp1_ds_mean - m_comp1_ds_sd, m_comp1_ds_mean + m_comp1_ds_sd),
    strTol = m_comp1_st_mean   # hold ST at mean
  )
)

summary(ss_comp1_je_ds, infer = c(TRUE, TRUE))
pairs(ss_comp1_je_ds)
```

```{r simple slopes for JE x ST interaction}
ss_comp1_je_st <- emtrends(
  m1a_fit_5dcr_comp1,
  ~ strTol,
  var = "joyExp",
  at = list(
    strTol = c(m_comp1_st_mean - m_comp1_st_sd, m_comp1_st_mean + m_comp1_st_sd),
    depSen = m_comp1_ds_mean   # hold other moderator constant (or choose a meaningful value)
  )
)

summary(ss_comp1_je_st, infer = c(TRUE, TRUE))
pairs(ss_comp1_je_st)
```

# How does trait curiosity predict Creating?

```{r prep data for 5DCR ~ Creating}
m2_data_5dcr_offloading <- df_offloading %>%
  left_join(df_curiosity_with_covars, by = "studId") %>% 
  drop_na() %>% 
  select(-n_total.y) %>% 
  rename(n_total = n_total.x)

print(paste("no. of students for 5DCR ~ offloading:", length(unique(m2_data_5dcr_offloading$studId))))
```


```{r Does 5DCR predict creating?}
m2_fit_5dcr_offloading <- glm(
  cbind(n_creating, (n_total - n_creating)) 
  ~ joyExp * depSen * strTol + dataset,
  family = binomial,
  data = m2_data_5dcr_offloading)

summary(m2_fit_5dcr_offloading)
Anova(m2_fit_5dcr_offloading, type = 3)
confint.default(m2_fit_5dcr_offloading, level = 0.95)
```
```{r sensitivity check: denom is not-na}
m2_fit_5dcr_offloading_s1 <- glm(
  cbind(n_creating, (n_is_not_na - n_creating))  ~ joyExp * depSen * strTol  + dataset,
  family = binomial,
  data = m2_data_5dcr_offloading)

summary(m2_fit_5dcr_offloading_s1)
Anova(m2_fit_5dcr_offloading_s1, type = 3)
confint.default(m2_fit_5dcr_offloading_s1, level = 0.95)
```

Simple slopes for JE x DS interaction on creating
```{r simple slopes for JE x DS interaction on creating}
m_off_ds_mean  <- mean(m2_data_5dcr_offloading$depSen, na.rm = TRUE)
m_off_ds_sd <- sd(m2_data_5dcr_offloading$depSen, na.rm = TRUE)
m_off_st_mean  <- mean(m2_data_5dcr_offloading$strTol, na.rm = TRUE)

# Simple slopes of JE at DS = mean ± 1 SD
ss_offloading_je_ds <- emtrends(
  m2_fit_5dcr_offloading,
  ~ depSen,
  var = "joyExp",
  at = list(
    depSen = c(m_off_ds_mean - m_off_ds_sd, m_off_ds_mean + m_off_ds_sd),
    strTol = m_off_st_mean   # hold ST constant
  )
)

summary(ss_offloading_je_ds, infer = c(TRUE, TRUE))
pairs(ss_offloading_je_ds)
```
```{r other perspective for simple slopes for JE x DS interaction on creating}
m_off_je_mean  <- mean(m2_data_5dcr_offloading$joyExp, na.rm = TRUE)
m_off_je_sd <- sd(m2_data_5dcr_offloading$joyExp, na.rm = TRUE)
m_off_st_mean  <- mean(m2_data_5dcr_offloading$strTol, na.rm = TRUE)
m_off_st_sd <- sd(m2_data_5dcr_offloading$strTol, na.rm = TRUE)

ss_offloading_ds_je <- emtrends(
  m2_fit_5dcr_offloading,
  ~ joyExp,
  var = "depSen",
  at = list(
    joyExp = c(m_off_je_mean - m_off_je_sd, m_off_je_mean + m_off_je_sd),
    strTol = m_off_st_mean   # hold ST constant
  )
)
summary(ss_offloading_ds_je, infer = c(TRUE, TRUE))
pairs(ss_offloading_ds_je)
```
When JE is low, greater DS results in greater % Creating



Simple slopes for DS and ST on creating
```{r simple slopes for DS and ST on creating}
m_off_je_mean  <- mean(m2_data_5dcr_offloading$joyExp, na.rm = TRUE)
m_off_st_sd <- sd(m2_data_5dcr_offloading$strTol, na.rm = TRUE)
m_off_ds_mean  <- mean(m2_data_5dcr_offloading$depSen, na.rm = TRUE)
m_off_ds_sd <- sd(m2_data_5dcr_offloading$depSen, na.rm = TRUE)
m_off_je_sd <- sd(m2_data_5dcr_offloading$joyExp, na.rm = TRUE)

ss_offloading_ds_st <- emtrends(
  m2_fit_5dcr_offloading,
  ~ strTol,
  var = "depSen",
  at = list(
    strTol = c(m_off_st_mean - m_off_st_sd, m_off_st_mean + m_off_st_sd),
    joyExp = m_off_je_mean   # hold JE constant
  )
)

summary(ss_offloading_ds_st, infer = c(TRUE, TRUE))
pairs(ss_offloading_ds_st)

```
When ST is high, DS is associated with greater Creating

```{r simple slopes for three way interaction}
m2_3way <- m2_fit_5dcr_offloading_s1

# Means and SDs of the moderators
m2_je_mean <- mean(m2_data_5dcr_offloading$joyExp, na.rm = TRUE)
m2_je_sd   <- sd(m2_data_5dcr_offloading$joyExp, na.rm = TRUE)

m2_st_mean <- mean(m2_data_5dcr_offloading$strTol, na.rm = TRUE)
m2_st_sd   <- sd(m2_data_5dcr_offloading$strTol, na.rm = TRUE)

# Simple slopes of depSen at low/high joyExp and low/high strTol
ss_depSen_3way <- emmeans::emtrends(
  m2_3way,
  ~ joyExp * strTol,
  var = "depSen",
  at = list(
    joyExp = c(m2_je_mean - m2_je_sd,
               m2_je_mean + m2_je_sd),
    strTol = c(m2_st_mean - m2_st_sd,
               m2_st_mean + m2_st_sd)
  )
)

summary(ss_depSen_3way, infer = c(TRUE, TRUE))

```


# Academics ~ 5DCR 

```{r Does 5DCR predict academic outcomes with imputation?}
m3_data_5dcr_imputAcads <- df_curiosity_with_covars %>%
  left_join(data_academic_outcomes_imputed %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z), by = "studId") %>%
  drop_na()
print(paste("no. of students for 5DCR ~ academics:", length(unique(m3_data_5dcr_imputAcads$studId))))

m3_fit_acad_5dcr <- lm(post_tgaia_z ~ pre_tgaia_z + depSen * joyExp * strTol + c_log_total + dataset, data = m3_data_5dcr_imputAcads)

summary(m3_fit_acad_5dcr)
Anova(m3_fit_acad_5dcr, type = 3)
```

```{r Does 5DCR predict academic outcomes without imputation?}
m3_data_5dcr_nonimputAcads <- df_curiosity_with_covars %>%
  left_join(data_academics_outcomes_nonimputed %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z), by = "studId") %>%
  drop_na()

print(paste("no. of students for 5DCR ~ academics (non-imputed):", length(unique(m3_data_5dcr_nonimputAcads$studId))))

m3_fit_acad_5dcr_nonimput <- lm(post_tgaia_z ~ pre_tgaia_z + depSen * joyExp * strTol + c_log_total + dataset, data = m3_data_5dcr_nonimputAcads)
summary(m3_fit_acad_5dcr_nonimput)
Anova(m3_fit_acad_5dcr_nonimput, type = 3)

```


# Academics ~ Prompting behaviours
```{r Do PCs predict academic outcomes with imputation?}
m4_data_pc_acads <- df_pca_ready %>%
  dplyr::select(studId, Comp1, Comp2) %>%
  left_join(data_academic_outcomes_imputed %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z), by = "studId") %>%
  left_join(df_covariates %>% dplyr::select(studId, dataset, c_log_total), by = "studId") %>%
  drop_na()
print(paste("no. of students for PCs ~ academics:", length(unique(m4_data_pc_acads$studId))))

m4_fit_acad_pc <- lm(post_tgaia_z ~ pre_tgaia_z + Comp1 * Comp2 + c_log_total + dataset, data = m4_data_pc_acads)

summary(m4_fit_acad_pc)
Anova(m4_fit_acad_pc, type = 3)
confint.default(m4_fit_acad_pc, level = 0.95)
effectsize::eta_squared(car::Anova(m4_fit_acad_pc, type = 3), partial = TRUE)
```
```{r Do PCs predict academic outcomes without imputation?}
m4_data_pc_acads_nonimput <- df_pca_ready %>%
  dplyr::select(studId, Comp1, Comp2) %>%
  left_join(data_academics_outcomes_nonimputed %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z), by = "studId") %>%
  left_join(df_covariates %>% dplyr::select(studId, dataset, c_log_total), by = "studId") %>%
  drop_na()

m4_data_pc_acads_nonimput <- df_pca_ready %>%
  dplyr::select(studId, Comp1, Comp2) %>%
  left_join(data_academics_outcomes_nonimputed %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z), by = "studId") %>%
  left_join(df_covariates %>% dplyr::select(studId, dataset, c_log_total), by = "studId") %>%
  drop_na()

print(paste("no. of students for PCs ~ academics (non-imputed):", length(unique(m4_data_pc_acads_nonimput$studId))))

m4_fit_acad_pc_nonimput <- lm(post_tgaia_z ~ pre_tgaia_z + Comp1*Comp2 + c_log_total + dataset, data = m4_data_pc_acads_nonimput)
summary(m4_fit_acad_pc_nonimput)
Anova(m4_fit_acad_pc_nonimput, type = 3)

```


## Creating and academic outcomes
```{r Does Create predict academic outcomes with imputation?}
m5_data_offloading_acads <- df_offloading %>%
  dplyr::select(studId, c_logit_off) %>%
  left_join(data_academic_outcomes_imputed %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z), by = "studId") %>%
  left_join(df_covariates %>% dplyr::select(studId, dataset, c_log_total), by = "studId") %>%
  drop_na()
print(paste("no. of students for offloading ~ academics:", length(unique(m5_data_offloading_acads$studId))))

m5_fit_acad_offloading <-  lm(post_tgaia_z ~ pre_tgaia_z + c_logit_off + c_log_total + dataset, data = m5_data_offloading_acads)
summary(m5_fit_acad_offloading)
Anova(m5_fit_acad_offloading, type = 3)
```
```{r Does Create predict academic outcomes without imputation?}
m5_data_offloading_acads_nonimput <- df_offloading %>%
  dplyr::select(studId, c_logit_off) %>%
  left_join(data_academics_outcomes_nonimputed %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z), by = "studId") %>%
  left_join(df_covariates %>% dplyr::select(studId, dataset, c_log_total), by = "studId") %>%
  drop_na()

print(paste("no. of students for offloading ~ academics (non-imputed):", length(unique(m5_data_offloading_acads_nonimput$studId))))

m5_fit_acad_offloading_nonimput <-  lm(post_tgaia_z ~ pre_tgaia_z + c_logit_off + c_log_total + dataset, data = m5_data_offloading_acads_nonimput)
summary(m5_fit_acad_offloading_nonimput)
Anova(m5_fit_acad_offloading_nonimput, type = 3)

```


