---
title: "Main analysis script for TeacherGAIA"
output: html_document
---
# README

This file comprises the analysis script for the TeacherGAIA project, using processed data from TeacherGAIA chatlogs. For privacy purposes, the raw student-chatbot exchanges may not be shared at this point.

This script is run on the files provided in ~/processed_data.

# Set Up
## Required packages
```{r Packages, message = FALSE, warning = FALSE}
pkgs = c("readxl", "car", "lme4", "lmerTest", "ggeffects", "psych", "tidyverse", "emmeans", "interactions", "patchwork", "svglite")

inst = lapply(pkgs, library, character.only=TRUE)

set.seed(555)
```
```{r Define file path}
root_dir <- '/Users/brainmemorylab/Library/CloudStorage/Dropbox-NUS/BrainMemLab_TeamFolder/Projects/Tidy_TeacherGAIA/05_expositions/code_clean/processed_data'  # Update this to your project root directory
```
```{r Colours for plots, include = FALSE}
colours_stress <- c("steelblue", "black", "royalblue4")
colours_joy <-c("darkgreen", "black", "darkolivegreen4")
colours_dep <-  c("goldenrod4","black", "darkorange2") 
```
```{r Set up contrasts for type III test of sig.}
options(contrasts = c("contr.sum", "contr.poly"))
```

# Import data
```{r}
# Mapping of studId to learning module topic (i.e., 'dataset')
data_studId_datasets <- readRDS(file.path(root_dir, "studId_dataset_mapping.rds"))

# Prompt-level labels across dimensions by Gemini 2.5 Pro, with confidence ratings and reasons 
data_query_intent <- read_xlsx(file.path(root_dir,'/agency_gem.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
    query_type = case_when(
      label == "Active Query" ~ 1L, 
      label == "Passive Response" ~ 0L,
      TRUE ~ NA_integer_
    )
  )

data_prompt_scope <- read_xlsx(file.path(root_dir,'/scope_gem.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
    prompt_scope = case_when(
      label == "Close-ended" ~ 1L,
      label == "Open-ended" ~ 2L, 
      label == "Not Applicable" ~ 3L,
      TRUE ~ NA_integer_
    )
  )

data_cognitive_intent <- read_xlsx(file.path(root_dir,'/intent_gem.xlsx'))  %>%
  mutate(label = case_when(label == "Irrelevant" ~ "Uncategorised", TRUE ~ label)) %>%
  mutate(
    cognitive_intent = case_when(
      label == "Remembering" ~ 1L,
      label == "Understanding" ~ 2L, 
      label == "Applying" ~ 3L, 
      label == "Analysing" ~ 4L,
      label == "Evaluating" ~ 5L,
      label == "Creating" ~ 6L, 
      label == "Uncategorised" ~ 7L,
      TRUE ~ NA_integer_
    )
  )  %>%
  mutate(
    studId = as.factor(studId),
  )

data_semantic_distance <- read_xlsx(file.path(root_dir,'/semantic_dist.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
    dataset = as.factor(dataset)
  )

# Academic outcomes: Imputed; keep 1st entry if duplicates present
df_student_level_academics_base <- read_xlsx(file.path(root_dir,'/replication_academics.xlsx')) %>%
  mutate(
    studId = as.factor(studId),
  )

# Processed 5DCR responses, wherein students who failed the attn check and straight-lined were dropped
df_student_level_curiosity <- read_xlsx(file.path(root_dir,'/replication_5dcr.xlsx')) %>%
  mutate(studId = as.factor(studId)) %>%
  drop_na(studId)
```

```{r Covariates preprocessing}
# df of studIds, 5DCR, z-test scores
df_student_covars <- data_studId_datasets %>%
  dplyr::select(studId, dataset) %>%
  mutate(dataset = droplevels(as.factor(dataset))) %>%
  left_join(
    df_student_level_curiosity %>% dplyr::select(studId, joyExp, depSen, strTol),
    by = "studId"
  ) %>%
  left_join(
    df_student_level_academics_base %>% dplyr::select(studId, pre_tgaia_z, post_tgaia_z),
    by = "studId"
  )
```

# Compute student averages of prompt features
```{r Logit function for proportions}
logit <- function(p) {
  p <- pmin(pmax(p, 1e-6), 1 - 1e-6)
  log(p / (1 - p))} 
```

```{r Prompt label preprocessing}
data_query_intent       <- data_query_intent       %>% mutate(id = as.character(id))
data_prompt_scope       <- data_prompt_scope       %>% mutate(id = as.character(id))
data_cognitive_intent   <- data_cognitive_intent   %>% mutate(id = as.character(id))
data_semantic_distance  <- data_semantic_distance  %>% mutate(id = as.character(id))

#1) Active vs. Passive Prompt
df_query_intent <- data_query_intent %>%
  inner_join(data_studId_datasets, by = "studId") %>%
  group_by(studId) %>%
  summarise(
    n_active_query      = sum(query_type == 1, na.rm = TRUE),
    n_passive_response  = sum(query_type == 0, na.rm = TRUE),
    n_total             = n_active_query + n_passive_response,
    dataset             = dplyr::first(dataset),
    .groups = "drop"
  ) %>%
  
  mutate(
    log_total = log1p(n_total),
    prop_active_adj = (n_active_query + 0.5) / (n_total + 1),
    logit_active = logit(prop_active_adj),
    c_log_total = as.numeric(scale(log_total, scale = TRUE)),
    c_logit_active = as.numeric(scale(logit_active, scale = TRUE))) %>% 
  inner_join(df_student_covars %>% dplyr::select(-dataset), by = "studId")
  

# 2) Open vs. Closed
df_prompt_scope <- data_prompt_scope %>%
  group_by(studId) %>%
  summarise(
    n_open  = sum(prompt_scope == 2, na.rm = TRUE),
    n_close = sum(prompt_scope == 1, na.rm = TRUE),
    n_na    = sum(prompt_scope == 3, na.rm = TRUE),
    n_openclose = n_open + n_close,
    n_total = n(),
    .groups = "drop"
  ) %>%
  mutate(
    prop_open_adj = (n_open + 0.5) / (n_openclose + 1),
    logit_open = logit(prop_open_adj),
    c_logit_open = as.numeric(scale(logit_open, scale = TRUE))
  ) %>%
  inner_join(df_student_covars, by = "studId")
  

# 3) Intent Categories
df_cognitive_intent <- data_cognitive_intent %>%
  mutate(
    is_high = as.integer(cognitive_intent %in% c(4,5,6)),
    is_high_nc = as.integer(cognitive_intent %in% c(4,5)),
    is_low  = as.integer(cognitive_intent %in% c(1,2,3)),
    is_create = as.integer(cognitive_intent %in% c(6)),
    is_na = as.integer(cognitive_intent == 7)
  ) %>%
  group_by(studId) %>%
  summarise(
    n_high = sum(is_high, na.rm = TRUE),
    n_high_nc = sum(is_high_nc, na.rm=TRUE),
    n_create = sum(is_create,na.rm=TRUE),
    n_low  = sum(is_low,  na.rm = TRUE),
    n_is_not_na = n_high + n_low,
    n_is_uncat = sum(is_na),
    n_total = n(),
      .groups = "drop"
  ) %>%
  mutate(
    prop_high_adj = (n_high + 0.5) / (n_is_not_na + 1),
    logit_high = logit(prop_high_adj),
    c_logit_high = as.numeric(scale(logit_high, scale = TRUE))
  ) %>%
  inner_join(df_student_covars, by = "studId")


# 4) Off-Loading
df_offloading <- data_cognitive_intent %>%
  group_by(studId) %>%
  summarise(
    n_creating  = sum(cognitive_intent == 6, na.rm = TRUE),
    n_is_not_na = sum(cognitive_intent < 7, na.rm = TRUE),
    n_total     = n(),
    .groups = "drop"
  ) %>%
  filter(n_is_not_na > 0) %>%
  inner_join(df_student_covars, by = "studId") %>%
  mutate(dataset = as.factor(dataset))

# 5) Semantic distance
df_student_level_semantics <- data_semantic_distance %>%
  group_by(studId) %>%
  summarise(n_total = n(),
            mean_p2p = mean(p2p, na.rm = TRUE),
            mean_sys = mean(sys)) %>% 
  mutate(mean_p2p_dist = 1-mean_p2p, # reverse-scored so higher scores reflect greater dissimilarity 
         mean_sys_dist = 1-mean_sys) %>% 
  dplyr::select(studId, mean_p2p_dist, mean_sys_dist, n_total) %>%  
  mutate(
    mean_p2p_dist_z = as.numeric(scale(mean_p2p_dist, scale = TRUE)),
    mean_sys_dist_z = as.numeric(scale(mean_sys_dist, scale = TRUE))
  ) %>% 
  inner_join(df_student_covars, by = "studId")

```

# Results
## 5DCR on interaction volume
```{r}
m_5dcr_volume <- lmer(c_log_total ~ joyExp * depSen * strTol + (1|dataset), data = df_query_intent)
summary(m_5dcr_volume)
```
Trait curiosity does not predict mere volume of prompts, which aligns with Jirout & Klahr (2012) and other work suggesting that curiosity is more about the quality of exploration rather than quantity.

## PCA 
```{r PCA}
# Prepare clean dataframe for PCA
df_pca_ready <- df_query_intent %>%
  dplyr::select(studId, c_logit_active, c_log_total, joyExp, depSen, strTol, dataset, pre_tgaia_z, post_tgaia_z) %>%
  
  left_join(df_prompt_scope %>% dplyr::select(studId, c_logit_open), by = "studId") %>%
  left_join(df_cognitive_intent %>%  dplyr::select(studId,c_logit_high), by = "studId") %>%
  left_join(df_offloading%>% dplyr::select(studId, c_logit_off), by = "studId") %>%
  left_join(df_student_level_semantics %>% dplyr::select(studId, mean_p2p_dist_z, mean_sys_dist_z), by = "studId") %>%
  drop_na(c_log_total, c_logit_active, c_logit_open, c_logit_high, mean_p2p_dist_z, mean_sys_dist_z)

# Check data suitability for PCA using KMO
pca_vars <- df_pca_ready %>%
  dplyr::select(c_logit_active, c_logit_open, c_logit_high, mean_p2p_dist_z, mean_sys_dist_z) %>%
  drop_na() %>% 
  as.data.frame()
kmo_result <- KMO(pca_vars)
print(kmo_result)
cortest.bartlett(pca_vars)

# Select and scale the 5 indicators
pca_input_clean <- df_pca_ready %>%
  dplyr::select(c_logit_active, c_logit_open, c_logit_high, mean_p2p_dist_z, mean_sys_dist_z) %>%
  as.data.frame()

# Parallel Analysis to select numbers of PC
parallel_results <- fa.parallel(pca_input_clean, 
                                fa = "pc", 
                                n.iter = 100, 
                                show.legend = TRUE, 
                                main = "Scree Plot with Parallel Analysis") # suggests 2

# Kaiser Criterion (Eigenvalues > 1.0)
actual_eigenvalues <- eigen(cor(pca_input_clean))$values
print("Actual Eigenvalues:")
print(round(actual_eigenvalues, 3)) # suggests 2

# Run the final 2-component model (based on Parallel analysis elbow)
final_pca_model <- principal(pca_input_clean, nfactors = 2, rotate = "varimax")
var_explained <- (final_pca_model$values / sum(final_pca_model$values)) * 100

print("Variance explained by each component:\n")
print(var_explained[1:2])
cat("Cumulative variance explained:", sum(var_explained[1:2]), "%\n")
print("PC loadings (cutoff = 0.3):\n")
print(final_pca_model$loadings, cutoff = 0.3)

# Extract scores back into dataframe
df_pca_ready$Comp1 <- final_pca_model$scores[, "RC1"]
df_pca_ready$Comp1 <- final_pca_model$scores[, "RC1"] * -1 # reverse score for interpretability 
df_pca_ready$Comp2 <- final_pca_model$scores[, "RC2"]
```
Parallel analysis and Kaiser both suggest 2 components. 2 Components explain 62.6% of beh variance.
PC1 is broadly associated with epistemic focus usage, while PC2 appears to reflect greater epistemic complexity. 

```{r Fig 1: PCA biplot}
loadings_df <- as.data.frame(final_pca_model$loadings[, 1:2])
loadings_df$RC1 <- loadings_df$RC1 * -1  # reverse PC1 loadings
loadings_df$variable <- rownames(loadings_df)

plot_pca_biplot <- ggplot(loadings_df, aes(x = RC1, y = RC2)) +
  geom_point(size = 6, color = "steelblue") +
  geom_text(aes(label = variable), nudge_x = 0.05, nudge_y = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_segment(aes(x = 0, y = 0, xend = RC1, yend = RC2),
               arrow = arrow(length = unit(0.3, "cm"))) +
  labs(x = "PC1 Loading",
       y = "PC2 Loading") +
  theme_classic() +
  theme(axis.text = element_text(size=12),
        axis.title = element_text(size=14))
print(plot_pca_biplot)
```

## Predicting PCA with 5DCR
```{r Model fits: Comp1 ~ 5DCR and Comp2 ~ 5DCR}
# Model 1: Predicting Component 1
m_comp1 <- lmer(Comp1 ~ depSen * joyExp * strTol + c_log_total + (1|dataset), data = df_pca_ready)
summary(m_comp1)
anova(m_comp1, type = 3)
confint(m_comp1)

# Model 2: Predicting Component 2
m_comp2 <- lmer(Comp2 ~ depSen * joyExp * strTol + c_log_total + (1|dataset), data = df_pca_ready)
summary(m_comp2)
anova(m_comp2, type = 3)
confint(m_comp2)
```
Interaction between DS x JE and JE x ST predict PC1

Follow-up analyses on the 2-way interactions
```{r Simple slopes for JE x DS on Comp 1}
ss_comp1 <- emtrends(m_comp1, ~ joyExp, var = "depSen", at = list(joyExp = c(-1, 1)))

summary(ss_comp1, infer = c(TRUE, TRUE))
pairs(ss_comp1)
```
Greater DS is associated with lower PC1 when JE is low, suggesting that when DS is high, people are less likely to engage in 'exploratory' type questioning.

```{r Fig 2A: DS x JE on PC1}
pred_PC1_JExDS <- ggpredict(m_comp1, terms = c("depSen", "joyExp [meansd]"))

lvl1 <- levels(factor(pred_PC1_JExDS$group))
pred_PC1_JExDS$group <- factor(
  pred_PC1_JExDS$group,
  levels = lvl1,
  labels = c("Low", "Mean", "High")
)

pred_PC1_JExDS <- pred_PC1_JExDS %>% 
  filter(group %in% c("Low", "High")) %>% 
  filter(!is.na(predicted))

fig2A_data <- df_pca_ready %>% filter(!is.na(Comp1), !is.na(depSen), !is.na(joyExp))

fig2A_PC1_JExDS <- 
  ggplot(pred_PC1_JExDS, aes(x = x, y = predicted, fill = group, color = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Low" = colours_dep[1], "High" = colours_dep[3])) +
  scale_fill_manual(values = c("Low" = colours_dep[1], "High" = colours_dep[3])) +
  labs(
    x = "Deprivation sensitivity",
    y = "Epistemic focus (PC1)",
    color = "Joyous exploration",
    fill = "Joyous exploration"
  ) +
  theme_classic() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.position = "bottom")

print(fig2A_PC1_JExDS)
```
```{r Simple slopes JE x ST on Comp 1}
ss_comp2 <- emtrends(m_comp1, ~ strTol, var = "joyExp", at = list(strTol = c(-1, 1)))

summary(ss_comp2, infer = c(TRUE, TRUE))
pairs(ss_comp2)
```
Greater JE is associated with more exploratory processing but only when ST is also high.

```{r Fig 2B: JE x ST on PC1}
pred_PC1_JExST <- ggpredict(m_comp1, terms = c("joyExp", "strTol [meansd]"))
lvl2 <- levels(factor(pred_PC1_JExST$group))
pred_PC1_JExST$group <- factor(
  pred_PC1_JExST$group,
  levels = lvl2,
  labels = c("Low", "Mean", "High")
)
pred_PC1_JExST <- pred_PC1_JExST %>% filter(group %in% c("Low", "High"))
fig2B_PC1_JExST <- 
  ggplot(pred_PC1_JExST, aes(x = x, y = predicted, fill = group, color = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Low" = colours_joy[1], "High" = colours_joy[3])) +
  scale_fill_manual(values = c("Low" = colours_joy[1], "High" = colours_joy[3])) +
  labs(
    x = "Joyous exploration",
    y = "Epistemic focus (PC1)",
    color = "Stress tolerance",
    fill = "Stress tolerance"
  ) +
  theme_classic() +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.position = "bottom")
print(fig2B_PC1_JExST)
```

## Predicting cognitive offloading with 5DCR
```{r Model fit: c_logit_off ~ 5DCR + n_total + (1|dataset)}
m_offloading <- glmer(
  cbind(n_creating, n_total - n_creating) ~
    joyExp * depSen * strTol + pre_tgaia_z + (1 | dataset),
  family = binomial,
  data = df_offloading
)
summary(m_offloading)
car::Anova(m_offloading)
confint(m_offloading, method = "Wald") # low incidence of Creating prompts
```
```{r Test the simple slopes for m4d_offloading_5dcr JE x DS interaction}
ss_offloading <- emtrends(m_offloading, ~ joyExp, var = "depSen", at = list(joyExp = c(-1, 1)))
summary(ss_offloading, infer = c(TRUE, TRUE))
pairs(ss_offloading)
```
```{r Plot the main effect of ST on offloading}
plot_offloading_mainST <- ggpredict(m_offloading, terms = "strTol") %>%
  plot(colors = colours_stress[1]) +
  labs(x = "Stress tolerance", y = "% Creating", title = NULL) +
  theme_classic() +
  geom_line(linewidth = 1.2, colour = colours_stress[1]) +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.position = "bottom")
print(plot_offloading_mainST)
```
```{r Plot of DS x JE on Off-loading}
pred_je_ds_off <-
  ggpredict(m_offloading, terms = c("depSen", "joyExp [meansd]"))
lvl7 <- levels(factor(pred_je_ds_off$group))
pred_je_ds_off$group <- factor(
  pred_je_ds_off$group,
  levels = lvl7,
  labels = c("Low", "Mean", "High")
)
# Keep only Mean-1SD and Mean+1SD
pred_je_ds_off <- pred_je_ds_off %>%
  filter(group %in% c("Low", "High"))

plot_offloading_JE_DS <- plot(pred_je_ds_off) +
  labs(x = "Deprivation sensitivity", y = "% Creating", title = NULL) +
  theme_classic() +
  scale_color_manual(name = "Joyous exploration", values = colours_dep[c(1, 3)]) +
  scale_fill_manual(name = "Joyous exploration", values = colours_dep[c(1, 3)]) +
  geom_line(linewidth = 1.2) +
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.position = "bottom")
print(plot_offloading_JE_DS)
```

## Predicting academic outcomes with information seeking profiles and metrics
### Components and academic outcomes
```{r Model fit: post_tgaia ~ pre_tgaia + Comp1*Comp2}
m_acad_pca <- lmer(post_tgaia_z ~ pre_tgaia_z + Comp1 * Comp2 * c_log_total + (1 | dataset), data = df_pca_ready, REML = FALSE)

summary(m_acad_pca)
anova(m_acad_pca,type = 3)
confint(m_acad_pca)
```
### Individual metrics and academic outcomes
```{r Model fit: post_tgaia ~ pre_tgaia + individual indices in the PCA + interaction with vol}
m_features_prompt_level <- lmer(post_tgaia_z ~ pre_tgaia_z + 
                                  # prompt features that were previously used to create the PCs
                                  (c_logit_active + c_logit_open + c_logit_high + mean_p2p_dist_z + mean_sys_dist_z)*c_log_total
                                + (1 | dataset), data = df_pca_ready, REML = FALSE)
summary(m_features_prompt_level)
anova(m_features_prompt_level, type = 3)
confint(m_features_prompt_level)
```
```{r FDR correction}
s <- summary(m_features_prompt_level)
pvals <- coef(s)[, "Pr(>|t|)"] # p-values for fixed effects
pvals_fdr <- p.adjust(pvals, method = "fdr")
pvals_fdr
```

### Offloading and academic outcomes
```{r Model fit: post_tgaia ~ pre_tgaia + c_logit_off + interaction with vol}
df_offloading <- df_offloading %>% left_join(df_query_intent %>% dplyr::select(studId, c_log_total), by = "studId")

m_offloading_acads <- lmer(post_tgaia_z ~ pre_tgaia_z + c_logit_off * c_log_total + (1 | dataset), data = df_offloading, REML = FALSE)

summary(m_offloading_acads)
anova(m_offloading_acads, type = 3)
```
