---
title: "Preprocessing of TeacherGAIA chatlogs"
output: 
  html_notebook:
    toc: TRUE
    toc_depth: 2
---
# README
This is the first step in preprocessing for the manuscript <i>"Trait curiosity shapes epistemic focus and exploration in naturalistic student-AI interactions"</i>.

A follow-up script ('tgaia_preprocessing.ipynb') is required to complete preprocessing and cleaning.

# Install requisite packages
```{r Install required packages}
pkgs = c("readxl", "writexl", "tidyEmoji", "cld2", "emoji", "tidyverse", "stringi")
inst = lapply(pkgs, library, character.only=TRUE)
set.seed(555)
```

# Data import
```{r}
root_dir <-'' # Our root directory
```

```{r Merge the two datafiles together}
t_raw1 <- read_xlsx(file.path(root_dir,'01_chatlogs.xlsx')) # File with raw student-chatbot exchanges

t_raw2 <- read_xlsx(file.path(root_dir,'02_systemprompts.xlxs')) # File with each chatbot's system prompt

data_tgaia_raw <- t_raw1 %>% full_join(t_raw2, by = "chatId") # Merge key is chatid
df_working <- data_tgaia_raw # Keep clean copy 
rm(t_raw1, t_raw2)
```

# Prompt-response cleaning
```{r Remove duplicate exchanges}
out_dupe_ids <- data_tgaia_raw[duplicated(data_tgaia_raw$id),]
df_working <- df_working %>% distinct(id, .keep_all = TRUE)
# removed 388 duplicated ids
```

```{r Remove observations where either prompt or response were missing}
out_missing_prompts <- df_working %>% filter(is.na(prompt))
out_missing_response <- df_working %>% filter(is.na(response))
df_working <- df_working %>% anti_join(out_missing_prompts, by = "id") %>% anti_join(out_missing_response, by = "id")
# removed 67 incomplete exchanges
```

# Manage systemPrompt and spId duplicates
Some learning modules were spread across multiple spIds, which resulted in duplicated systemPrompts. We identified these and merged them to create "class" that shared a learning module topic. 
```{r Identify the duplicated spIds}
t_dupe_spIds <- df_working %>% select(spId, systemPrompt) %>% 
  distinct(spId, .keep_all = TRUE) %>% 
  arrange(systemPrompt)
# identify duplicated systemPrompts
out_dupe_spIds <- t_dupe_spIds %>%
  group_by(systemPrompt) %>%
  filter(n() > 1) %>% 
  arrange(systemPrompt)
```

- Duplicated system prompts:
  - 284 & 285
  - 481 & 592
  - 314 & 291 & 283
  - 639 & 606
  - 638 & 605
  - 640 & 722

```{r Create new dataset labels}
df_working <- df_working %>%
  mutate(dataset = case_when(
    spId == 284 ~ "SG_5W1H",
    spId == 285 ~ "SG_5W1H",
    
    spId == 481 ~ "Contemporary_SG",
    spId == 592 ~ "Contemporary_SG",
    
    spId == 283 ~ "Singapore_Significance",
    spId == 291 ~ "Singapore_Significance",
    spId == 314 ~ "Singapore_Significance",
    
    spId == 639 ~ "AI_Aware",
    spId == 606 ~ "AI_Aware",
    spId == 605 ~ "AI_Aware",
    spId == 638 ~ "AI_Aware",
    
    spId == 640 ~ "Argumentative_Essay",
    spId == 722 ~ "Argumentative_Essay",
    
    spId == 365 ~ "Knowing_Korea",
    spId == 313 ~ "Vietnam_War",
    spId == 401 ~ "Self_Governance_SG", # later renamed to Singapore Significance 2
    spId == 407 ~ "Inclusive_Housing",
    spId == 545 ~ "Unpacking_Poems",
    spId == 756 ~ "History_Teacher",

    TRUE ~ as.character(spId) 
  ))
```

# Removal of any non-semantic prompts 
Non-semantic defined as a lack of text (i.e., alphanumeric characters) and ONLY an emoji
```{r Remove all non-semantic rows}
df_working <- df_working %>%
  mutate(detected.emoji = emoji_detect(df_working$prompt)) 

out_rows_with_emojis <- df_working %>% filter(detected.emoji == TRUE)
out_rows_with_emojis_ONLY <- out_rows_with_emojis %>% filter(!stri_detect_regex(prompt, "[A-Za-z]"))

df_working <- df_working %>% anti_join(out_rows_with_emojis_ONLY, by = "id")
```

# Managing non-English text
```{r Remove that Chinese agent}
df_working %>% filter(spId == 485) %>% count()
df_working <- df_working %>% filter(!spId == 485)
# removed 3,174 instances from spId 485, a Mandarin-conducted assignment
```

```{r Overwrite non-English text with translated values}
df_working$language <- detect_language(df_working$prompt)
out_non_english_rows <- df_working %>% filter(!language == "en")

# translations provided by Google Translation API
data_translations <- read.csv(file.path(root_dir,'/03_translations_checked.csv')) %>%
  filter(!spId == 485) %>% 
  select(id, prompt_translated, response_translated) %>% 
  distinct(id, .keep_all = TRUE)

df_working <- df_working %>% left_join(data_translations, by ="id")

df_working <- df_working %>% 
  mutate(prompt = if_else(!is.na(prompt_translated), prompt_translated, prompt)) %>% 
  select(-prompt_translated) %>% 
  mutate(response = if_else(!is.na(response_translated), response_translated, response)) %>% 
  select(-response_translated)
```

# Final processed dataset
```{r}
## split by spId
#   spId_split <- split(df_working, df_working$spId)
#   
#   for (spId in names(spId_split)) {
#     
#     filename <- file.path(file.path(root_dir,'/02_processed_data/output/split_by_spId', paste0(spId, ".xlsx")))
#     
#     openxlsx::write.xlsx(spId_split[[spId]], file = filename, rowNames = FALSE)
#   }
#   
## split by learning module label
#   dataset_split <- split(df_working, df_working$dataset)
#   
#   for (dataset in names(dataset_split)) {
#     
#     filename <- file.path(file.path(root_dir,'/02_processed_data/output/split_by_dataset', paste0(dataset, ".xlsx")))
#     
#     openxlsx::write.xlsx(dataset_split[[dataset]], file = filename, rowNames = FALSE)
#   }

# full compiled dataset
write_xlsx(df_working, file.path(root_dir,'/output/compiled_tgaia.xlsx'))
```

# Selection criteria
Selected for nine learning modules with knowledge acquisition goals
```{r Select learning modules}
selected_datasets <- c("Vietnam_War", "Self_Governance_SG", "Inclusive_Housing", "Contemporary_SG", "History_Teacher", "AI_Aware", "Singapore_Significance", "Knowing_Korea", "Unpacking_Poems")

df_dataset <- df_working %>% filter(dataset %in% selected_datasets)

write_xlsx(df_dataset,file.path(root_dir,'/output/selected_tgaia.xlsx'))
```

# Cleaning of excel artefacts from prompt and response text
Please refer to 'tgaia_preprocessing.ipynb' for the cleaning of excel artefacts from prompt and response text, and conversion to .pkl file. 
